---
title: "M5_Lab4_Time_series_in_R_crime_data_Cody_York"
output: html_notebook
---

**Import Libraries:** Also setting working directory.
```{r}
library(tidyr)
library(dbplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(ggfortify)
library(plotly)
library(forecast)

setwd('C:/Users/cy_su/PycharmProjects/DSCI_605_Data_Visualizations/Module 5/M5_Lab4/')
```

**Read the data from the CSV file:**
Convert the "Date" column to a datetime format
Filter the data for the desired two-year period.
```{r}
data <- read.csv("Fouryears_all.csv")

data$Date <- ymd_hms(data$Date)

start_date <- as.Date("2019-09-01")
end_date <- as.Date("2021-09-01")
filtered_data <- data %>%
  filter(Date >= start_date & Date <= end_date & Primary.Type == "BATTERY")
```

**Data Wrangling:**
Aggregate the data by hourly intervals.
Display the aggregated data.
```{r}
data_hourly <- filtered_data %>%
  mutate(Hour = floor_date(Date, "hour")) %>%
  group_by(Hour) %>%
  summarise(Count = n()) %>%
  ungroup()

```

**Time Series Visualization**
Create a time series object.
Plot the time series.
Show the plot.
Save the image.

What I did to improve the visualization:
    1. Changed Line Color and Type.
    2. Added point density.
    3. Legend and Title Positions
    4. Added a color palette
```{r}
plot <- ggplot(data = data_hourly, aes(x = Hour, y = Count)) +
  geom_smooth(aes(color = "Count"), se = FALSE, size = 1.2, linetype = "solid") +
  geom_point(size = 2, color = "blue", alpha = 0.5) +
  scale_x_datetime(date_labels = "%b %d, %Y", date_breaks = "1 day") +
  scale_color_brewer(palette = "Set1", name = "Metrics") +
  labs(x = "Date (Year-Month-Day)",
       y = "Count",
       title = "2-Year Series of Battery Crimes In California",
       subtitle = "Hourly Data Aggregated by Day",
       caption = "Source: Chicago Police Department’s CLEAR (Citizen Law Enforcement Analysis and Reporting) System") +
  theme_minimal() +
  theme(axis.text = element_text(size = 10, color = "darkblue"),
        axis.title = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face = "italic", hjust = 0.5),
        plot.caption = element_text(size = 8, hjust = 1),
        legend.position = "bottom",
        panel.grid.major = element_line(color = "gray", size = 0.1),
        panel.grid.minor = element_line(color = "gray", size = 0.05),
        plot.background = element_rect(fill = "aliceblue")) +
  guides(color = guide_legend(override.aes = list(size = 4)))

ggsave("ts_plot.png")

plotly_plot <- ggplotly(plot)

plotly_plot


```

**Decompose the Time Series:**
Convert decomposed object to a data frame.
Plot the data using autoplot.
Save the image.
```{r}
data_ts <- ts(data_hourly$Count, frequency = 24) # Adjust frequency as needed

data_ts_decomposed <- decompose(data_ts)

autoplot(data_ts_decomposed)

ggsave("ts_decomposed_plot.png")
```


**Create a MSTS Object:**
Decompose the object.
Plot the object.
Save the image.
```{r}
battery_msts <- msts(data_hourly$Count, seasonal.periods = c(24,24*7,24*30))

battery_msts_decompose <- mstl(battery_msts)

autoplot(battery_msts_decompose)

ggsave("ts_msts_plot.png")
```
